---
title: "The EntiCe study"
author: "S. Rao"
date: "5/8/2020"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R Markdown

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Introduction-1

The EntiCe study, published in 2019, was carried out by a UPenn team consisting of basic researchers, veternarians and CHOP pediatricians led by Prof. Daniel Beiting. 

The driving hypothesis: inflammatory bowel disease in dogs (a model for the human version) is likely a result of an altered gut microbiome. 

By restoring this microbiome to a healthy one using a special diet of hydrolysed protein, diet-responsive dogs showed a marked improvement in their symptoms.

## Introduction-2  

  The study tracked gut microbiome changes in 29 dogs exhibiting IBD symptoms over the course of 42 days. 
  
  The primary means of tracking was NGS based, 16S rRNA metabarcoding of stool samples taken from these dogs along with appropriate healthy controls. 
  
## Introduction-3

  The study was a comprehensive test of the hypothesis and included a novel metabolomic identification of putative antibiotics that restore the healthy microbiome. 
  
  We can't recommend enough that you read the original open access paper to understand the very comprehensive approach taken by the researchers. Here is a link to the paper:
  
  https://link.springer.com/article/10.1186/s40168-019-0740-4
  
## Introduction-4

  Here, I focus on a re-analysis of the metabarcoding using Qiime2 and R.
  
## Slide with bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

